---
title: "design-overview"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{design-overview}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(evolveTrial)
```

## Overview

`evolveTrial` provides tools for designing and simulating Bayesian adaptive
clinical trials with time-to-event or binary endpoints. This vignette walks
through a short end-to-end workflow:

1. Find a binary two-stage (Simon) design and check its exact operating
   characteristics.
2. Evaluate a single-arm time-to-event design by simulation.
3. Configure and simulate a hybrid seamless single-arm-to-between-arm design.

Computationally heavy simulation chunks are shown with `eval = FALSE` so the
vignette builds quickly; copy them into an interactive session to run them.

## 1. Binary two-stage (Simon) design

`find_simon_design()` searches for an optimal or minimax Simon two-stage design,
and `simon_oc_exact()` returns its exact operating characteristics. These are
fast, closed-form computations, so we evaluate them here.

```{r simon}
set.seed(2026)

# Optimal Simon design: H0: p = 0.20 vs H1: p = 0.40
design <- find_simon_design(p0 = 0.20, p1 = 0.40, alpha = 0.05, beta = 0.20)
design

# Exact operating characteristics under the null (p = 0.20) and the
# alternative (p = 0.40)
simon_oc_exact(
  n1 = design$n1, r1 = design$r1,
  n  = design$n,  r  = design$r,
  p  = 0.20
)
simon_oc_exact(
  n1 = design$n1, r1 = design$r1,
  n  = design$n,  r  = design$r,
  p  = 0.40
)
```

## 2. Single-arm time-to-event design

`run_simulation_pure()` simulates a single-arm trial with a piecewise
exponential time-to-event model and Bayesian posterior decision rules, returning
type I error / power, probabilities of early termination, and expected sample
size. This is a Monte Carlo simulation, so we mark it `eval = FALSE`.

```{r single-arm, eval = FALSE}
set.seed(2026)

result <- run_simulation_pure(
  num_simulations            = 500,
  arm_names                  = "Experimental",
  reference_arm_name         = "Experimental",
  compare_arms_option        = "none",
  weibull_shape_true_arms    = c(Experimental = 1),
  weibull_median_true_arms   = c(Experimental = 9),
  interval_cutpoints_sim     = c(0, 3, 6, 9, 12, 15, 18, 21, 24),
  max_follow_up_sim          = 24,
  censor_max_time_sim        = 24,
  prior_alpha_params_model   = rep(0.01, 8),
  prior_beta_params_model    = rep(0.01, 8),
  num_posterior_draws        = 5000,
  cohort_size_per_arm        = c(Experimental = 1),
  max_total_patients_per_arm = c(Experimental = 60),
  efficacy_stopping_rule_hc  = TRUE,
  efficacy_threshold_hc_prob = 0.95,
  futility_stopping_rule_hc  = TRUE,
  futility_threshold_hc_prob = 0.05,
  median_pfs_success_threshold_arms = c(Experimental = 6),
  overall_accrual_rate       = 2,
  randomization_probs        = c(Experimental = 1),
  min_events_hc              = 15,
  interim_calendar_beat      = 3
)

print(result)
```

## 3. Hybrid seamless single-arm-to-between-arm design

A hybrid design begins with single-arm monitoring and can convert to a
between-arm comparison once sufficient evidence accrues. `create_hybrid_theta()`
builds the design parameter vector and `compute_hybrid_oc_rcpp()` simulates the
operating characteristics via the C++ engine. Again `eval = FALSE` because this
runs many Monte Carlo replications.

```{r hybrid, eval = FALSE}
theta <- create_hybrid_theta(
  eff_sa             = 0.90,
  fut_sa             = 0.10,
  hr_threshold_sa    = 0.80,
  ev_sa              = 15,
  nmax_sa            = 40,
  conversion_trigger = "any_single_success",
  pp_go              = 0.70,
  pp_nogo            = 0.20,
  ss_method          = "predictive",
  max_additional_n   = 60,
  eff_ba             = 0.975,
  fut_ba             = 0.05,
  ev_ba              = 15,
  nmax_ba            = 80
)

base_args <- list(
  interval_cutpoints_sim   = c(0, 3, 6, 9, 12, 15, 18, 21, 24),
  prior_alpha_params_model = rep(0.01, 8),
  prior_beta_params_model  = rep(0.01, 8),
  overall_accrual_rate     = 3,
  interim_calendar_beat    = 3,
  num_posterior_draws      = 5000
)

# Scenario: historical median 6 mo, reference 6 mo, experimental 9 mo
scenario <- list(historical_median = 6, ref_median = 6, exp_median = 9)

oc <- compute_hybrid_oc_rcpp(
  hybrid_theta    = theta,
  base_args       = base_args,
  scenario_params = scenario,
  num_simulations = 1000,
  seed            = 42,
  trial_mode      = "hybrid"
)

cat("Power:", round(oc$power, 3), "\n")
cat("Type I:", round(oc$type1, 3), "\n")
cat("E[N] under null:", round(oc$EN_null, 1), "\n")
cat("P(conversion):", round(oc$P_conversion, 3), "\n")
```

## Where to go next

See the package README for the full API reference (simulation, calibration,
binary-endpoint helpers, visualization, and decision-rule functions) and the
companion `BATON` package for constrained Bayesian optimization over the design
space evaluated here.
